From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-05-29 paper.tex
Cleanups

High-fidelity display systems are the integration point of the Blue Brain
visualization infrastructure. They are the evolution of existing visualization
systems, enabling higher resolution, immersion and team collaboration. Compared
to current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a simple to use environment for high-fidelity visualization.
For all use cases, the increased display size and resolution allows improved
data exploration for 2D and 3D content, facilitating large data exploration.

@@

The core of the Blue Brain visualization infrastructure are multiple
high-resolution tiled display walls driven by our Tide software~\cite{tide}. All
walls are equipped with a multitouch user interface and can be remote controlled
from any web browser. The walls are build using thin-bezel, 55 inch, Full HD LCD
panels with a hardened glass sheet. We use $4\times 3$ and $3\times 3$
configurations for a total of 24 and 18 Megapixel resolution, respectively. The
display size of over five meter diagonally (four meter for $3\times 3$) allows
team-size collaboration (up to ten people) or project-wide presentations (up to
a hundred people). \fig{fTDW} shows one wall during a project-wide presentation
with multiple interactive applications running concurrently on the wall.

@@

In the last years we implemented a production-ready setup of monoscopic tiled
display walls. On one hand, while there is substantial research on tiled display
wall software~\cite{Sage, Sage2, Omegalib, DK:11, JLMV:06, DisplayCluster}, we
found that most solutions where not ready to be used in a $24\times 7$
unattained environment. For this reason, we started with the TACC DisplayCluster
open source software~\cite{DisplayCluster}, which we incrementally refactored
and improved to the current TIDE implementation.

On the other hand, the technology has been commoditized to make these type of
installations affordable to medium-sized institutions which allowed us to build
the software integration for a reasonable startup cost. We have focused on the
multitouch user interface, which implements a low entry barrier for new users,
a unique capability of our solution.

@@

OpenDeck is our next-generation visualization system, aiming to integrate the
success of tiled display walls with a seamless transition to fully immersive
environments. We are currently in the process of installing the system which
consists of a semi-cylindrical back-projection screen with 41 Megapixel usable
resolution on a 36 $m^2$ surface (\fig{fOpenDeck}). Like the display walls, it
is equipped with multitouch capabilities which makes it usable as a monoscopic
collaboration system from the first day of installation. Unlike tiled display
walls, it is active stereo capable and equipped with a 3D tracking system for
immersive rendering. For increased immersion, a lower resolution front
projection system fills in the floor area. OpenDeck will allow us to run our
applications based on Equalizer~\cite{EMP:09} once the system is installed.

@@

The OpenDeck will provide a unique environment for the research and development
of new visualization techniques. The OpenDeck immersive infrastructure with
multitouch will open a set of questions along immersive touch user interfaces,
transitions and mixing of monoscopic to immersive usage, the combination of
tracked and touch devices, multi-user immersion, latency for remote immersive
rendering as well as multi-site collaboration.

@@

Compared to other solutions~\cite{Sage,Sage2,DisplayCluster}, the development of
Tide was driven by the need of a stable and easy to use implementation. Tide
supports three types of content: files (high-resolution images, movies, pdfs),
built-in applications (web browser, whiteboard) and remote applications using
the Deflect library (DesktopStreamer, Equalizer-based applications, Brayns).
The multitouch user interface can handle multiple users manipulating different
windows and their content.

@@

Deflect is the client library for Tide. It provides an API for pixel streaming
to Tide and for receiving events from Tide. The pixel streaming allows
synchronized parallel streaming from a parallel rendering application as well as
monoscopic and stereoscopic streams. Various events allow the application to
react to multi-touch input from the wall.

@@

@@

All Blue Brain applications integrate a messaging libraries which allow them
to be used as services in a variety of use cases. For example, the Tide web
server providing the user interface shown in \fig{fTideWeb}, is based on this
messaging solution. Other use cases are remote python APIs, JavaScript user
interfaces and service architectures combining multiple visualization
applications with data providers such as HPC simulations (\fig{fZero}).

@@

ZeroEQ is the base messaging library, wrapping up existing technologies into an
API which is convenient to use and easy to integrate into C++ code. It provides
two messaging services: publish-subscribe and HTTP. For binary and JSON
encodings it relies on a simple \textsf{Serializable} interface, for which
ZeroBuf provides a sample implementation. To facilitate the simple use case of
linking a few applications, ZeroEQ uses the zeroconf protocol to discover and
connect to related applications. For more complex scenarios, explicit
connections are supported.

@@

Advances in computer hardware and software have brought raytracing to the point
where it replaces classical rasterization for virtually all use cases in
scientific visualization. On one hand, local illumination for typical OpenGL
data sets (up to hundred million triangles) can be done at similar framerates
to OpenGL~\cite{ospray}. For small datasets, OpenGL performs better, but for
larger data sets raytracing outperforms OpenGL. This is due to a better
scalability with respect to the data set size ($O(\log{N})$ vs $O(n)$).
Furthermore, CPU-based raytracing allows the rendering of larger data sets
without any level-of-detail algorithms due to the larger memory size. Last, but
not least, advanced rendering algorithms such as shadowing, reflections and
global illuminations are significantly easier to implement in a raytracer. The
only area where rasterization provides a benefit is for rendering at very high
frame rates, needed for example for immersive visualization.

It is for these reasons that interactive raytracing is the future technology for
interactive and offline rendering at Blue Brain. We have developed a first open
source implementation of a visualization engine with different backends:
OSPRay~\cite{ospray} for CPU-based raytracing and OptiX~\cite{optix} for
GPU-based raytracing. These application called Brayns can load and visualize
our data sets in a variety of modes and integrates with our messaging solution
to be accessible from Python. We are currently integrating the key algorithms
from Livre for out-of-core volume rendering in Brayns.

>>>>

High-fidelity display systems are the integration point of the Blue Brain
visualization capabilities. They are the evolution of existing visualization
systems, enabling higher resolution, immersion and team collaboration. Compared
to current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a more approachable environment for high-fidelity
visualization. For all use cases, the increased display size and resolution
allows better data exploration for 2D and 3D content, facilitating large data
exploration.

@@

The core of the Blue Brain visualization infrastructure are multiple
high-resolution tiled display walls driven by our Tide software~\cite{tide}. All
walls are equipped with a multitouch user interface and can be remote controlled
from any web browser. The walls are build using thin-bezel, 55 inch, Full HD LCD
panels with a hardened glass sheet. We use $4\times 3$ and $3\times 3$
configurations for a total of 24 and 18 Megapixel resolution, respectively. The
display size of over five meter diagonally (four meter for $3\times 3$) allows
team-size collaboration (up to ten people) or project-wide presentations (up to
a hundred people). \fig{fTDW} shows one wall during a project-wide presentation
with multiple interactive applications running remotely on the wall.

@@

While there is substantial research on tiled display wall software~\cite{Sage,
Sage2, Omegalib, DK:11, JLMV:06, DisplayCluster}, we found that most solutions
where not ready production use in a $24\times 7$ unattained environment. For
this reason, we started with the TACC DisplayCluster open source
software~\cite{DisplayCluster}, which we incrementally refactored and improved
to the current TIDE implementation. On the other hand, the hardware has been
commoditized to make these type of installations affordable to medium-sized
institutions which allowed us to build the software integration for a reasonable
startup cost. We have focused on the multitouch user interface, which implements
a low entry barrier for new users, a unique capability of our solution.

@@

OpenDeck is our next-generation visualization system, aiming to integrate the
success of tiled display walls with a seamless transition to fully immersive
environments. We are currently in the process of installing a system which
consists of a semi-cylindrical back-projection screen with 41 Megapixel usable
resolution on a 36 $m^2$ surface (\fig{fOpenDeck}). Like the display walls, it
is equipped with multitouch capabilities which makes it usable as a monoscopic
collaboration system from the first day of installation. Unlike tiled display
walls, it is active stereo capable and equipped with a 3D tracking system for
immersive rendering. For increased immersion, a lower resolution front
projection system fills in the floor area. OpenDeck will run TIDA and our
immersive applications based on Equalizer~\cite{EMP:09} once the system is
installed.

@@

OpenDeck will provide a unique environment for the research and development of
new visualization techniques. It will open a set of questions along immersive
touch user interfaces, transitions and mixing of monoscopic to immersive usage,
the combination of tracked and touch devices, multi-user immersion, latency for
remote immersive rendering as well as multi-site collaboration.

@@

Compared to other solutions~\cite{Sage,Sage2,DisplayCluster}, the development of
Tide was driven by the need of a stable and easy to use implementation. Tide
supports three types of content: files (high-resolution images, movies, pdfs),
built-in applications (web browser, whiteboard) and remote applications using
the Deflect library (DesktopStreamer, Equalizer-based applications, Brayns).
The multitouch user interface can handle multiple users manipulating different
windows and their content simultaneously.

@@

Deflect is the client library for Tide. It provides an API for pixel streaming
to Tide and for receiving events from Tide. The pixel streaming allows
synchronized parallel streaming from parallel rendering application as well as
monoscopic and stereoscopic streams. Various events allow the application to
react to multi-touch input from the wall.

@@

Advances in computer hardware and software have brought raytracing to the point
where it replaces classical rasterization for virtually all use cases in
scientific visualization. On one hand, OpenGL-like local illumination for
typical data sets used with rasterization (up to hundred million triangles) can
be done at similar framerates to OpenGL~\cite{ospray}. For small datasets,
OpenGL performs better, but for larger data sets raytracing outperforms OpenGL.
This is due to a better scalability with respect to the data set size
($O(\log{N})$ vs $O(n)$). Furthermore, CPU-based raytracing allows the rendering
of larger data sets without any level-of-detail algorithms due to the larger
memory size. Last, but not least, advanced rendering algorithms such as
shadowing, reflections and global illuminations are significantly easier to
implement in a raytracer. The only area where rasterization provides a benefit
is for rendering at very high frame rates, needed for example for immersive
visualization.

It is for these reasons that interactive raytracing is the future technology for
interactive and offline rendering at Blue Brain. We have developed a first open
source implementation of a visualization engine with different backends:
OSPRay~\cite{ospray} for CPU-based raytracing and OptiX~\cite{optix} for
GPU-based raytracing. This application called Brayns can load and visualize our
data sets in a variety of modes and integrates with our messaging solution to be
accessible from Python. We are currently integrating the key algorithms from
Livre for out-of-core volume rendering in Brayns, which will also facilitate
mixing polygonal with volume data in a single scene.

@@

All Blue Brain applications integrate messaging libraries which allow them to be
used as services in a variety of use cases. For example, the Tide web server
providing the user interface shown in \fig{fTideWeb}, is based on this messaging
solution. Other use cases are remote python APIs, JavaScript user interfaces and
service architectures combining multiple visualization applications with data
providers such as HPC simulations (\fig{fZero}).

@@

ZeroEQ is our C++ messaging library, wrapping up existing technologies into an
API which is convenient to use and easy to integrate into C++ code. It provides
two messaging services: publish-subscribe and HTTP. For binary and JSON
encodings it relies on a simple \textsf{Serializable} interface, for which
ZeroBuf provides a sample implementation. To facilitate the simple use case of
linking a few applications, ZeroEQ uses the zeroconf protocol to discover and
connect to related applications. For more complex scenarios, explicit
connections are supported.

@@
